From clusters to queries: exploiting uncertainty in the modularity landscape of complex networks

### 2018-06-25 article/article.tex
Rewrite of lfr tests to us new cross validation method

\section{Results}
\subsection{Cross-validation method}
\label{sec:cross_validation}
In this study we test a small number of labels that we intend to use in order to evaluate how well our method correctly generalises to discover unlabelled vertices.
We would like to test a significantly smaller number of seed nodes than two class classification methods used in previous studies, which use leave-one-out cross validation \cite{kohler2008walking}.
The cross validation procedure we devise is described as follows and depends on the size of the community and the number of initial seed labels being used.
For this work we would like to capture binary classification  performance, \textit{true positives (tp)}, \textit{true negatives (tn), false positives (fp)} and \textit{false negatives (fn)}, on our datasets of community memberships.
In order to generate the different sets for the cross validation we take each label set and generate unique sample sets of vertices from the known true positive labels.
We also note that although each node can have multiple labels, in the tests performed in this study we only consider performance on individual labels.

@@

\subsection{Random walk with restart}
As a comparison, we compare our method with the commonly used random walk with restart method as described in \cite{kohler2008walking}.
This simulates an infinite random walk with a probability, $\alpha$ that a walker would teleport back to the initial seed nodes. 
Formally, the random walk with restart algorithm uses the form
\begin{equation} \label{eq:rwr}
 \vec{p}_{t+1} = (1 - \alpha) W \vec{p}_t + \alpha \vec{p}_0,
\end{equation}
where $W$ is the row normalised adjacency matrix of a graph, $\vec{p}_0$ is the initial walk vector such that $\vec{p}_0 = \frac{1}{|S|}$ for vertices in the seed set.
The random walk algorithm repeats equation \ref{eq:rwr} from $\vec{p}_0$ until the L1 norm between  $\vec{p}_t$ and  $\vec{p}_{t+1}$ converges to 0.
For the tests performed in this work we use a restart probability $\alpha = 0.25$.

@@

\subsubsection{Tests on gene ontology labels}
\begin{figure*}[]
     \centering
    \begin{subfigure}[b]{0.31\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/rocs/yeast_ppi_amigo.eps}
        \caption{Yeast PPI GO terms}
    \end{subfigure}
    \begin{subfigure}[b]{0.31\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/rocs/arabidopsis_ppi_amigo.eps}
        \caption{Arabidopsis PPI GO terms}
    \end{subfigure}
    \begin{subfigure}[b]{0.31\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/rocs/ecoli_ppi_amigo.eps}
        \caption{E. coli PPI GO terms}
    \end{subfigure}
    \caption{Receiver operator characteristic (ROC) curves for gene ontology labels for real networks described in Section \ref{sec:real_networks}. 
    Results represent the averages over all receiver operating characteristic scores from all samples of label sets with the cross validation procedure described in \ref{sec:cross_validation}.}
    \label{fig:GO_TERM_ROCS}
\end{figure*}

\begin{figure*}[t]
\includegraphics[width=0.8\textwidth]{images/ppi_example_combined.eps}
\caption{An example of the GO term \textit{``SNAP receptor activity''} (16 nodes) on the \textit{Arabidopsis thaliana} protein-protein interaction network.
(A) Shows the full network with 3 randomly chosen query nodes from the full query set.
(B) Shows a reduced view of the network where a threshold of $\mu_i(S) > 0.8$ is selected, reducing the network to 404 nodes.
Blue nodes indicates the 3 randomly selected query nodes, yellow the remaining nodes in the term and the shade of red relates to the strength of the relationship each node has to the query set.
(C) Shows an example ROC curve from the cross validation described in the main text.
}
\label{fig:query_example}
\end{figure*}

@@

>>>>

\section{Results}
\subsection{Cross-validation method}
\label{sec:cross_validation}
In this study we test a small number of labels that we intend to use in order to evaluate how well our method correctly generalises to discover unlabelled vertices.
We would like to test a significantly smaller number of seed nodes than two class classification methods used in previous studies, which use leave-one-out cross validation \cite{kohler2008walking}.
The cross validation procedure we devise is described as follows and depends on the size of the community and the number of initial seed labels being used.
For this work we would like to capture binary classification  performance, \textit{true positives (tp)}, \textit{true negatives (tn), false positives (fp)} and \textit{false negatives (fn)}, on our datasets of community memberships.
In order to generate the different sets for the cross validation we take each label set and generate unique sample sets of vertices from the known true positive labels\footnote{Note that although each node can have multiple labels, in the tests performed in this study we only consider performance on individual labels as this is likely the situation one would encounter with limited label sets in practice.}.

@@

In the case for the synthetic networks tested below, true labels are drawn from the known communities, with each tested in isolation.
For the real world networks tested the labels are considered in the same manner though many nodes have no assigned labels.
Every community is considered with different seed set sizes.

\subsection{Random walk with restart}
In the following sections compare our method with the commonly used random walk with restart method as described in \cite{kohler2008walking}.
This simulates an infinite random walk with a fixed probability, $\alpha$ that a walker would teleport back to the initial seed nodes. 
Formally, the random walk with restart algorithm uses the form
\begin{equation} \label{eq:rwr}
 \vec{p}_{t+1} = (1 - \alpha) W \vec{p}_t + \alpha \vec{p}_0,
\end{equation}
where $W$ is the row normalised adjacency matrix of a graph, $\vec{p}_0$ is the initial walk vector such that $\vec{p}_0 = \frac{1}{|S|}$ for vertices in the seed set.
The random walk algorithm repeats equation \ref{eq:rwr} from $\vec{p}_0$ until the L1 norm between  $\vec{p}_t$ and  $\vec{p}_{t+1}$ converges to 0, simulating the steady state vector $\vec{p}_{\infty}$.
For the tests performed in this work we use a restart probability $\alpha = 0.25$.

@@

\subsubsection{Tests on gene ontology labels}
\begin{figure*}[t]
     \centering
    \begin{subfigure}[b]{0.31\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/rocs/yeast_ppi_amigo.eps}
        \caption{Yeast PPI GO terms}
    \end{subfigure}
    \begin{subfigure}[b]{0.31\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/rocs/arabidopsis_ppi_amigo.eps}
        \caption{Arabidopsis PPI GO terms}
    \end{subfigure}
    \begin{subfigure}[b]{0.31\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/rocs/ecoli_ppi_amigo.eps}
        \caption{E. coli PPI GO terms}
    \end{subfigure}
    \caption{Receiver operator characteristic (ROC) curves for gene ontology labels for real networks described in Section \ref{sec:real_networks}. 
    Results represent the averages over all receiver operating characteristic scores from all samples of label sets with the cross validation procedure described in \ref{sec:cross_validation}.}
    \label{fig:GO_TERM_ROCS}
\end{figure*}

\begin{figure}[t]
\includegraphics[width=0.48\textwidth]{images/ppi_example_combined.eps}
\caption{An example of the GO term \textit{``SNAP receptor activity''} (16 nodes) on the \textit{Arabidopsis thaliana} protein-protein interaction network.
(A) Shows the full network with 3 randomly chosen query nodes from the full query set.
(B) Shows a reduced view of the network where a threshold of $\mu_i(S) > 0.8$ is selected, reducing the network to 404 nodes.
Blue nodes indicates the 3 randomly selected query nodes, yellow the remaining nodes in the term and the shade of red relates to the strength of the relationship each node has to the query set.
(C) Shows an example ROC curve from the cross validation described in the main text.
}
\label{fig:query_example}
\end{figure}

@@

The reader may also consider the similarity of this approach to that of label propagation \cite{raghavan2007near, gregory2010finding}, which seeks to find communities based on a vote, where at different time steps a node updates it's label to to be the most common amongst its neighbours.
A common problem with label propagation schemes is that they often fail to converge, resulting in many competing clusterings of a network.
We liken this to the problem encountered by Good Good \textit{et al.} \cite{good2010performance} in modularity maximisation.
The core contribution of this paper is that, in such situations, there is no single, context independent labelling scheme that can be seen as the ``true'' community structure, overlapping or not.
Indeed, the approach applied here is extremely general and could be adapted to the label propagation approach (or any other community detection algorithm) should sufficient semi-supervised group memberships be known \textit{a priori}.
